### 2015-06-11 doc/abc.tex
first

\documentclass[preprint]{aastex}
\usepackage{amsmath,amssymb}
\usepackage{mathrsfs}

\bibliographystyle{plain}

\begin{document}

\title{Type Ia Supernova Model}
\author{Alex Kim}

\section{Science Interest}
The discovery of the accelerating expansion of the Universe came as a great surprise,
although such behavior was already known to be a theoretical possibility.
General Relativity (GR) long provided a basis for such an acceleration through
the Cosmological Constant $\Lambda$, which as a constant vacuum energy density has the
negative pressure that leads to gravitational repulsion.  An expectation
for vacuum energy density also comes from particle physics but naive predictions anticipate
a higher energy density than that observed: the existence of a
non-zero but small Cosmological Constant is a major challenge confronting physics theory.

There are no viable first-principles physical models, rather the acceleration caused by
``dark energy''  with an equation of state of fairly unconstrained functional form. The parameterized model used to analyze data are based on a Taylor expansion for which GR is the zeroth-order solution.

Most physicists think that the accelerating universe is due to
GR's $\Lambda$. The ``easiest'' great physics discovery that can be made with the next 
generation of surveys is to establish the inconsistency of the data relative to the
expectations of GR; only if the data can exclude GR do we have 
a hope to track down the nature of the next generation of physics.  Therefore, it seems that 
the first layer of statistical analysis should be hypothesis testing with
GR as the null hypothesis.

The technique we are familiar with is the goodness-of-fit test using $p$-values.
The likelihood does not reduce to a simple least-squares
so a $\chi^2$-distribution does not apply.  The model of the data involves latent parameters
drawn from unknown distributions, and objects not discovered due to survey inefficiency.

GR makes a prediction for the distance modulus as a function of redshift,
\begin{equation}
\mu_\Lambda(z;H_0, \Omega_M, \Omega_\Lambda),
\end{equation}
where  three free parameters are accommodated, or two when flatness is assumed i.e.\
$\Omega_M+\Omega_\Lambda=1$.  For the purposes of hypothesis testing,
$H_0$, $\Omega_M$, $\Omega_\Lambda$ are nuisance parameters.

Each object (transient) in the Universe has intrinsic properties
\begin{equation}
\{M(\lambda,t), A(\lambda,t), z, \mu\},
\end{equation}
where $M$ is the absolute magnitude, $z$ represents the cosmological redshift (peculiar velocities are ignored), and $\mu$ is the distance modulus corresponding to that redshift.

We access information about these objects
\begin{equation}
\{\hat{T}, \hat{m}(t), \hat{H}, \hat{z}, \hat{\mu}
\},
\end{equation}
where $\hat{T}$ is a boolean whether the transient is discovered and $\hat{m}$ are the measurements.
$\hat{H}$ is the weight of the object in the analysis, clearly non-discovery implies $\hat{H}=0$
and traditional Type~Ia SN classification gives a binary $\hat{H} \in \{0,1\}$ for consideration in
the cosmology analysis.  The measured redshift $\hat{z}$ can have uncertainty due to host-misidentification and further spectroscopic or photometric redshift uncertainty.  $\hat{mu}$
is a number usually derived from the use of an SN~Ia model to interpret $\hat{m}$.

The underlying function $\mu(z)$ is what we want to infer from the data.  Of particular interest
is to show that $\mu(z)$ is inconsistent with the theoretical expectation from the theoretical expectation for a Cosmological Constant universe.

\end{document}